---
title: "integrative analysis of inflamed/non-inflamed intestinal tissues of 6 GEO datasets"
author: Peinan Zhao
---

```{r, include=FALSE}
library(GEOquery)
library(edgeR)
library(limma)
library(clusterProfiler)
library(ReactomePA)
library(org.Hs.eg.db)
library(enrichplot)
library(readxl)
library(UpSetR)
library(org.Hs.eg.db)
library(CEMiTool)
library(WGCNA)
library(msigdbr)
library(affy)
library(DescTools)
library(VennDiagram)
library(RColorBrewer)
library(xlsx)
library(tidyverse)

#load("analysis_result.Rdata")
```

## Results {.tabset}

### load datasets

```{r, include=FALSE}
include.study <- c("38713","95437","179285","59071","75214","83687")
DE_files <- list.files("DE_genes", full.names = T)

DE.genes.list <- list()
active.genes.list <- list()
remission.genes.list <- list()

active.genes.list.1 <- list()
remission.genes.list.1 <- list()

for(study in include.study){
  DE.table <- read.csv(DE_files[grepl(study, DE_files)]) 
  
  DE.genes.list[[study]] <- dplyr::filter(DE.table,!duplicated(DE.table$Gene))$Gene
  remission.genes.list[[study]] <- dplyr::filter(DE.table, logFC>0 & !duplicated(DE.table$Gene))$Gene
  active.genes.list[[study]] <- dplyr::filter(DE.table, logFC<0 & !duplicated(DE.table$Gene))$Gene

  remission.genes.list.1[[study]] <- tibble(gene = remission.genes.list[[study]], !!study := rep(1, length(remission.genes.list[[study]])))
  active.genes.list.1[[study]] <- tibble(gene = active.genes.list[[study]], !!study := rep(1, length(active.genes.list[[study]])))
}

remission.matrix <- remission.genes.list.1 %>%
  purrr::reduce(full_join, by = "gene") %>%
  replace(is.na(.), 0) %>%
  mutate(sum = rowSums(across(where(is.numeric))))

active.matrix <- active.genes.list.1 %>%
  purrr::reduce(full_join, by = "gene") %>%
  replace(is.na(.), 0) %>%
  mutate(sum = rowSums(across(where(is.numeric))))


remission.matrix.sort <- remission.matrix[order(-remission.matrix$sum),]
#write_csv(remission.matrix.sort, "tables/common_remission_genes.csv")

active.matrix.sort <- active.matrix[order(-active.matrix$sum),]
#write_csv(active.matrix.sort, "tables/common_active_genes.csv")
```

### UpSetR

```{r, fig.width=10, fig.height=5, echo=FALSE}
include.study <- paste("GSE", include.study, sep = "")
comset <- CombSet(include.study, 5, repl = F, ord = F)
ints <- list()
queries <- list
for (i in c(1:length(comset[,1]))){
  ints[[i]] <- c(comset[i,])
}
ints[[i+1]] <- c(include.study)

up.plot <- as.data.frame(remission.matrix[,c(2:7)])
colnames(up.plot) <- paste("GSE",colnames(up.plot), sep = "")
#pdf("figures/UpSetR_remission.pdf", width = 10, height = 5)
upset(up.plot, mainbar.y.label = "Remission marker genes",nsets = 6, order.by = c("freq"),
      queries = list(
        list(query = intersects, params = ints[[1]], color = "orange", active = T),
        list(query = intersects, params = ints[[5]], color = "orange", active = T),
        list(query = intersects, params = ints[[6]], color = "orange", active = T),
        list(query = intersects, params = ints[[7]], color = "orange", active = T)), 
      point.size = 2, number.angles = 20)
#dev.off()

#pdf("figures/UpSetR_active.pdf", width = 10, height = 5)
up.plot <- as.data.frame(active.matrix[,c(2:7)])
colnames(up.plot) <- paste("GSE",colnames(up.plot), sep = "")
upset(up.plot, mainbar.y.label = "Active marker genes",nsets = 6, order.by = c("freq"),
      queries = list(
        list(query = intersects, params = ints[[1]], color = "orange", active = T),
        list(query = intersects, params = ints[[5]], color = "orange", active = T),
        list(query = intersects, params = ints[[6]], color = "orange", active = T),
        list(query = intersects, params = ints[[7]], color = "orange", active = T)), 
      point.size = 2, number.angles = 20)
#dev.off()

```

Common Remission marker genes: Up regulation in remission group in **>= 5** studies

```{r, message=FALSE, warning=FALSE}
common.remission <- dplyr::filter(remission.matrix, sum >= 5)$gene
common.remission.Entrez <- mapIds(org.Hs.eg.db, keys = common.remission, column = "ENTREZID", keytype = "SYMBOL")

GO.common.remission <- enrichGO(gene        = common.remission.Entrez,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05, readable = T)
GO.common.remission <- GO.common.remission@result
GO.common.remission$Fold.enrichment <- calculateFoldEnrichment(GO.common.remission)
GO.common.remission <- GO.common.remission  %>%
  filter(p.adjust < 0.01)

#write.xlsx(GO.common.remission, file = "figures/common_DEGs_GOBP.xlsx", sheetName = "Remission", row.names = F, append = F)
```

Common Active marker genes: Up regulation in remission group in **>= 5** studies

```{r}
common.active <- dplyr::filter(active.matrix, sum >=5)$gene
common.active.Entrez <- mapIds(org.Hs.eg.db, keys = common.active, column = "ENTREZID", keytype = "SYMBOL")

GO.common.active <- enrichGO(gene        = common.active.Entrez,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05, readable = T)
GO.common.active <- GO.common.active@result
GO.common.active$Fold.enrichment <- calculateFoldEnrichment(GO.common.active)
GO.common.active <- GO.common.active  %>%
  filter(p.adjust < 0.01)

#write.xlsx(GO.common.active, file = "figures/common_DEGs_GOBP.xlsx", sheetName = "Inflamed", row.names = F, append = T)
```

all DEGs

```{r}
all.DEGs <- rbind(remission.matrix, active.matrix)
at_least_3 <- filter(all.DEGs, sum >=3)
at_least_4 <- filter(all.DEGs, sum >=4)
at_least_5 <- filter(all.DEGs, sum >=5)
at_least_6 <- filter(all.DEGs, sum >=6)
```

UpsetR plot for DE genes in >= 5 studies

```{r}
# combined
up.plot1 <- as.data.frame(remission.matrix) %>%
  rbind(as.data.frame(active.matrix)) %>%
  filter(sum >= 5) %>%
  mutate(type = ifelse(gene %in% common.remission, "2", "3"))
colnames(up.plot1) <- paste("GSE",colnames(up.plot1), sep = "")

p <- upset(up.plot1, mainbar.y.label = "DE genes",nsets = 15, order.by = c("freq"), point.size = 2, 
      number.angles = 0, main.bar.color = "#EC6914", show.numbers = F, query.legend = "top", text.scale = 1.3,
      queries = list(
        list(query = elements, params = list("GSEtype", c(2,3)), color = "#EC6914", active = T, 
             query.name = "Active"),
        list(query = elements, params = list("GSEtype", 2), color = "#044c94", active = T, 
             query.name = "Remission")
      ))
pdf("figures/UpSetR_combined.pdf", width = 5.5, height = 5)
p
dev.off()

#remission
up.plot2 <- as.data.frame(remission.matrix) %>%
  filter(sum >= 5)
colnames(up.plot2) <- paste("GSE",colnames(up.plot2), sep = "")
p <- upset(up.plot2, mainbar.y.label = "Common remission genes",nsets = 15, order.by = c("freq"), point.size = 2,
           text.scale = 1.3, number.angles = 0, main.bar.color = "#044c94")
pdf("figures/UpSetR_remission.pdf", width = 6, height = 5)
p
dev.off()

#active
up.plot3 <- as.data.frame(active.matrix) %>%
  filter(sum >= 5)
colnames(up.plot3) <- paste("GSE",colnames(up.plot3), sep = "")
p <- upset(up.plot3, mainbar.y.label = "Common active genes",nsets = 15, order.by = c("freq"), point.size = 2, 
      number.angles = 0, main.bar.color = "#EC6914", text.scale = 1.3)
pdf("figures/UpSetR_active.pdf", width = 5, height = 5)
p
dev.off()
```

### GO analysis

```{r, include=FALSE}
calculateFoldEnrichment <- function(GO){
  df <- data.frame(GeneRatio = GO$GeneRatio, BgRatio = GO$BgRatio)
  df[] <- apply(df, c(1, 2), function(x) eval(parse(text = x)))
  fold.enrich <- df$GeneRatio/df$BgRatio
  return(fold.enrich)
}
```

Find over-represented GO-BP genesets by remission/non-inflamed samples in each GEO dataset

```{r, include=FALSE}
remission.gene.EntrezID <- as.data.frame(remission.matrix)
remission.gene.EntrezID$EntrezID <- mapIds(org.Hs.eg.db, keys = remission.gene.EntrezID$gene, keytype = "SYMBOL", column = "ENTREZID")

## 38713
GO.38713.remission <- enrichGO(gene          = filter(remission.gene.EntrezID, `38713` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.38713.remission <- GO.38713.remission@result
GO.38713.remission$Fold.enrichment <- calculateFoldEnrichment(GO.38713.remission)
GO.38713.remission.sig <- GO.38713.remission %>%
  filter(p.adjust < 0.01)

### 95437
GO.95437.remission <- enrichGO(gene          = filter(remission.gene.EntrezID, `95437` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.95437.remission <- GO.95437.remission@result
GO.95437.remission$Fold.enrichment <- calculateFoldEnrichment(GO.95437.remission)
GO.95437.remission.sig <- GO.95437.remission  %>%
  filter(p.adjust < 0.01)


### 179285
GO.179285.remission <- enrichGO(gene          = filter(remission.gene.EntrezID, `179285` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.179285.remission <- GO.179285.remission@result
GO.179285.remission$Fold.enrichment <- calculateFoldEnrichment(GO.179285.remission)
GO.179285.remission.sig <- GO.179285.remission  %>%
  filter(p.adjust < 0.01)


### 59071
GO.59071.remission <- enrichGO(gene          = filter(remission.gene.EntrezID, `59071` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.59071.remission <- GO.59071.remission@result
GO.59071.remission$Fold.enrichment <- calculateFoldEnrichment(GO.59071.remission)
GO.59071.remission.sig <- GO.59071.remission  %>%
  filter(p.adjust < 0.01)

### 75214
GO.75214.remission <- enrichGO(gene          = filter(remission.gene.EntrezID, `75214` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.75214.remission <- GO.75214.remission@result
GO.75214.remission$Fold.enrichment <- calculateFoldEnrichment(GO.75214.remission)
GO.75214.remission.sig <- GO.75214.remission  %>%
  filter(p.adjust < 0.01)

### 83687
GO.83687.remission <- enrichGO(gene          = filter(remission.gene.EntrezID, `83687` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.83687.remission <- GO.83687.remission@result
GO.83687.remission$Fold.enrichment <- calculateFoldEnrichment(GO.83687.remission)
GO.83687.remission.sig <- GO.83687.remission  %>%
  filter(p.adjust < 0.01)

remission.sig.GO <- list(GSE38713 = GO.38713.remission.sig, 
                         GSE95437 = GO.95437.remission.sig, 
                         GSE179285 = GO.179285.remission.sig, 
                         GSE59071 = GO.59071.remission.sig, 
                         GSE75214 = GO.75214.remission.sig, 
                         GSE83687 = GO.83687.remission.sig)
remission.GO.list <- list()
for(study in names(remission.sig.GO)){
  remission.GO.list[[study]] <- tibble(GO = remission.sig.GO[[study]]$Description, 
                                       !!study := rep(1, length(remission.sig.GO[[study]]$ID)))
}

remission.GO.matrix <- remission.GO.list %>%
  purrr::reduce(full_join, by = "GO") %>%
  replace(is.na(.), 0) %>%
  mutate(ave = rowSums(across(where(is.numeric)))/6) %>%
  arrange(-ave)
remission.GO.df <- as.data.frame(remission.GO.matrix)[c(2:7)]

remission.GO.term <- remission.GO.matrix$GO[1:20]
remission.GO.all <- do.call("rbind", list( GSE95437 = GO.95437.remission.sig, 
                                           GSE179285 = GO.179285.remission.sig, 
                                           GSE59071 = GO.59071.remission.sig, 
                                           GSE75214 = GO.75214.remission.sig, 
                                           GSE83687 = GO.83687.remission.sig,
                                           GSE38713 = GO.38713.remission.sig)) %>%
  filter(Description %in% remission.GO.term)

remission.GO.all$Dataset <- gsub(".GO.*","",rownames(remission.GO.all))
remission.GO.all$Dataset <- factor(remission.GO.all$Dataset,
                                   levels = unique(remission.GO.all$Dataset))

write.csv(remission.GO.all, "tables/Remission_GO_all.csv")
```

Find over-represented GO-BP genesets by inflamed samples in each GEO dataset

```{r, include=FALSE}
active.gene.EntrezID <- as.data.frame(active.matrix)
active.gene.EntrezID$EntrezID <- mapIds(org.Hs.eg.db, keys = active.gene.EntrezID$gene, keytype = "SYMBOL", column = "ENTREZID")

## 38713
GO.38713.active <- enrichGO(gene          = filter(active.gene.EntrezID, `38713` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.38713.active <- GO.38713.active@result
GO.38713.active$Fold.enrichment <- calculateFoldEnrichment(GO.38713.active)
GO.38713.active.sig <- GO.38713.active %>%
  filter(p.adjust < 0.01)

### 95437
GO.95437.active <- enrichGO(gene          = filter(active.gene.EntrezID, `95437` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.95437.active <- GO.95437.active@result
GO.95437.active$Fold.enrichment <- calculateFoldEnrichment(GO.95437.active)
GO.95437.active.sig <- GO.95437.active  %>%
  filter(p.adjust < 0.01)


### 179285
GO.179285.active <- enrichGO(gene          = filter(active.gene.EntrezID, `179285` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.179285.active <- GO.179285.active@result
GO.179285.active$Fold.enrichment <- calculateFoldEnrichment(GO.179285.active)
GO.179285.active.sig <- GO.179285.active  %>%
  filter(p.adjust < 0.01)


### 59071
GO.59071.active <- enrichGO(gene          = filter(active.gene.EntrezID, `59071` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.59071.active <- GO.59071.active@result
GO.59071.active$Fold.enrichment <- calculateFoldEnrichment(GO.59071.active)
GO.59071.active.sig <- GO.59071.active  %>%
  filter(p.adjust < 0.01)

### 75214
GO.75214.active <- enrichGO(gene          = filter(active.gene.EntrezID, `75214` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.75214.active <- GO.75214.active@result
GO.75214.active$Fold.enrichment <- calculateFoldEnrichment(GO.75214.active)
GO.75214.active.sig <- GO.75214.active  %>%
  filter(p.adjust < 0.01)

### 83687
GO.83687.active <- enrichGO(gene          = filter(active.gene.EntrezID, `83687` == 1)$EntrezID,
                               OrgDb         = org.Hs.eg.db,
                               ont           = "BP",
                               pAdjustMethod = "BH",
                               pvalueCutoff  = 0.01,
                               qvalueCutoff  = 0.05)
GO.83687.active <- GO.83687.active@result
GO.83687.active$Fold.enrichment <- calculateFoldEnrichment(GO.83687.active)
GO.83687.active.sig <- GO.83687.active  %>%
  filter(p.adjust < 0.01)

active.sig.GO <- list(GSE38713 = GO.38713.active.sig, 
                         GSE95437 = GO.95437.active.sig, 
                         GSE179285 = GO.179285.active.sig, 
                         GSE59071 = GO.59071.active.sig, 
                         GSE75214 = GO.75214.active.sig, 
                         GSE83687 = GO.83687.active.sig)
active.GO.list <- list()
for(study in names(active.sig.GO)){
  active.GO.list[[study]] <- tibble(GO = active.sig.GO[[study]]$Description, 
                                       !!study := rep(1, length(active.sig.GO[[study]]$ID)))
}

active.GO.list <- list()
for(study in names(active.sig.GO)){
  active.GO.list[[study]] <- tibble(GO = active.sig.GO[[study]]$Description, 
                                       !!study := -log10(active.sig.GO[[study]]$p.adjust))
}

active.GO.matrix <- active.GO.list %>%
  purrr::reduce(full_join, by = "GO") %>%
  replace(is.na(.), 0) %>%
  mutate(ave = rowSums(across(where(is.numeric)))/6) %>%
  arrange(-ave)
active.GO.df <- as.data.frame(active.GO.matrix)[c(2:7)]

active.GO.term <- active.GO.matrix$GO[1:20]
active.GO.all <- do.call("rbind", list( GSE95437 = GO.95437.active.sig, 
                                           GSE179285 = GO.179285.active.sig, 
                                           GSE59071 = GO.59071.active.sig, 
                                           GSE75214 = GO.75214.active.sig, 
                                           GSE83687 = GO.83687.active.sig,
                                           GSE38713 = GO.38713.active.sig)) %>%
  filter(Description %in% active.GO.term)

active.GO.all$Dataset <- gsub(".GO.*","",rownames(active.GO.all))
active.GO.all$Dataset <- factor(active.GO.all$Dataset,
                                   levels = unique(active.GO.all$Dataset))
active.GO.all$Description <- factor(active.GO.all$Description,
                                   levels = rev(active.GO.term))
write.csv(active.GO.all, "tables/active_GO_all.csv")
```

Combined GO figure

```{r, fig.height=8, fig.width=15}
DEG.list.entrez <- list()
for (geo in names(remission.genes.list)){
  EntrezID <- mapIds(org.Hs.eg.db, keys = remission.genes.list[[geo]], keytype = "SYMBOL", column = "ENTREZID")
  EntrezID <- EntrezID[!is.na(EntrezID)]
  DEG.list.entrez[[paste("GSE", geo, "_remission", sep = "")]] <- EntrezID
}

for (geo in names(active.genes.list)){
  EntrezID <- mapIds(org.Hs.eg.db, keys = active.genes.list[[geo]], keytype = "SYMBOL", column = "ENTREZID")
  EntrezID <- EntrezID[!is.na(EntrezID)]
  DEG.list.entrez[[paste("GSE", geo, "_active", sep = "")]] <- EntrezID
}

combinedGO <- compareCluster(geneClusters = DEG.list.entrez, fun = "enrichGO", OrgDb = org.Hs.eg.db, ont = "BP", pvalueCutoff = 0.01, qvalueCutoff = 0.01)

pdf("figures/GO_enrichment_combined.pdf", width = 13, height = 7)
dotplot(combinedGO, showCategory = 3, by="geneRatio") + xlab("") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 11, face = "bold"), axis.title.y = element_text(size = 12))
dev.off()

combinedGO.read <- setReadable(combinedGO, OrgDb = org.Hs.eg.db)

combinedGO.table <- combinedGO.read@compareClusterResult %>%
  mutate(Fold.enrichment = calculateFoldEnrichment(combinedGO.read@compareClusterResult))
write.csv(combinedGO.table, "figures/GEO_dataset_GOBP_all.csv")
```

### CEMiTool

Load data from GSE179285

```{r, include=FALSE}
gset <- getGEO("GSE179285", GSEMatrix =TRUE, AnnotGPL=TRUE)
if (length(gset) > 1) idx <- grep("GPL6480", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

# make proper column names to match toptable
fvarLabels(gset) <- make.names(fvarLabels(gset))

# Get sample names in each group
metaData <- pData(phenoData(gset))

active <- row.names(dplyr::filter(metaData,`inflammation:ch1`=="Inflamed"))
remission <- row.names(dplyr::filter(metaData,`inflammation:ch1`=="Uninflamed" & `diagnosis:ch1`!="Healthy control"))
sel <- c(active, remission)

group <- factor(c(rep("active", length(active)),rep("remission", length(remission))))

## Filter, keep the samples in the comparison groups
gset <- gset[ ,sel]
exprs(gset) <- normalizeBetweenArrays(exprs(gset)) # normalize data - between array

ex <- exprs(gset)
gset$group <- factor(group)
metaData <- pData(phenoData(gset))
```

Generate count matrix using common DE genes

```{r, message=FALSE, warning=FALSE}
common.label <- dplyr::filter(gset@featureData@data, Gene.symbol %in% c(common.remission, common.active)) %>%
  dplyr::filter(!duplicated(Gene.symbol))

expr <- ex[rownames(common.label),]
expr <- as_data_frame(log2(expr))
rownames(expr) <- common.label$Gene.symbol

sample.annot <- tibble(SampleName = rownames(gset@phenoData@data), Class=gset@phenoData@data$group)
sample.annot <- as.data.frame(sample.annot)
```

Analysis

```{r, message=FALSE, warning=FALSE}
cem <- new_cem(expr = expr, sample_annot = sample.annot, filter = F)
cem <- find_modules(cem, cor_method = "pearson", network_type = "signed")

cem <- plot_beta_r2(cem)
pdf("figures/CEMiTool/Soft_threshold.pdf", width = 7, height = 5)
cem@beta_r2_plot
dev.off()

cem <- plot_qq(cem)
cem@qq_plot
```

GSEA analysis

```{r, message=FALSE, warning=FALSE}
cem <- mod_gsea(cem)
cem <- plot_gsea(cem)
pdf("figures/CEMiTool/Module_enrichment.pdf", width = 7, height = 5)
show_plot(cem, "gsea")
dev.off()
```

GO_biological process enrichment

```{r, message=FALSE, warning=FALSE}
bp.genesets.all <- msigdbr(species = "Homo sapiens", category = "C5", subcategory = "BP")
subsets.bp <- data_frame(term=bp.genesets.all$gs_name, gene=bp.genesets.all$human_gene_symbol)
subsets.bp$term <- gsub("GOBP_","",subsets.bp$term)

cem <- mod_ora(cem, subsets.bp)
cem <- plot_ora(cem, pv_cut=0.05)
write_csv(cem@ora, "tables/Gene_modules_GO_enrichment.csv")

cem@barplot_ora$M5
cem@barplot_ora$M1

cem@barplot_ora$M2
cem@barplot_ora$M3
cem@barplot_ora$M4


pdf("figures/CEMiTool/GO_M5.pdf", width = 7, height = 5)
cem@barplot_ora$M5
dev.off()

pdf("figures/CEMiTool/GO_M1.pdf", width = 7, height = 5)
cem@barplot_ora$M1
dev.off()

pdf("figures/CEMiTool/GO_M2.pdf", width = 7, height = 5)
cem@barplot_ora$M2
dev.off()

pdf("figures/CEMiTool/GO_M3.pdf", width = 7, height = 5)
cem@barplot_ora$M3
dev.off()

pdf("figures/CEMiTool/GO_M4.pdf", width = 7, height = 5)
cem@barplot_ora$M4
dev.off()
```

Get hub genes

```{r}
GOI <- c()
hubs <- get_hubs(cem, n = 30, method = "adjacency")
for (i in c(1:length(hubs))){
  GOI <- c(GOI, names(hubs[[i]]))
}
GOI <- unique(GOI)
remission.hubgenes <- c(hubs$M2, hubs$M3, hubs$M4)
```

Adding protein interaction network for GOI (protein interaction data are retrived from string database version.11 https://string-db.org/)


```{r}
# GOI are the hub genes of gene modules
GOI <- c()
hubs <- get_hubs(cem, n = 30, method = "adjacency")
for (i in c(1:length(hubs))){
  GOI <- c(GOI, names(hubs[[i]]))
}
GOI <- unique(GOI)

# read files downloaded from string database
protein_db <- read.table(file = "data/9606.protein.links.v11.0.txt",header = TRUE, stringsAsFactors = F)
string_proteins <- read.table(file = "data/human.name_2_string.tsv", header = FALSE, stringsAsFactors = F)

# get protein interaction pairs of GOI
get_PPI_pairs <- function(protein_db, string_proteins, min_combined_score, GOI){
  GOI_1 <- GOI
  find.GOI.1 <- string_proteins %>%
    filter(V2 %in% GOI_1)  
  GOI_1.id <- find.GOI.1$V3
  GOI_1 <- find.GOI.1$V2
  int.df.1 <- tibble(gene1symbol=0, gene2symbol=0)
  for (i in 1:length(GOI_1)){
    count = 0
    gene2symbol = c()
    int.temp <- protein_db %>%
      filter(combined_score > min_combined_score) %>% # only retrieve high confident interactions
      filter(protein1 == GOI_1.id[i])
    if (length(int.temp$protein1)){
      for (n in 1:length(int.temp$protein2)) {
          find.protein <- string_proteins %>%
            filter(V3 == int.temp$protein2[n])
          if(length(find.protein$V2)){
            gene2symbol[n] <- find.protein$V2
          }
        }
        int.df.1 <- rbind(int.df.1,tibble(gene1symbol = rep(GOI_1[i], length(gene2symbol)), gene2symbol = gene2symbol))
    }
  }
    int.df.1 <- as.data.frame(int.df.1) %>%
      filter(gene1symbol != 0)%>%
      filter(!is.na(gene1symbol) & !is.na(gene2symbol))
    return(int.df.1)
}

int.df <- get_PPI_pairs(protein_db = protein_db, string_proteins = string_proteins, min_combined_score = 700, GOI = GOI)
```

Add protein interactions to analysis

```{r}
interactions_data(cem) <- int.df.1 # add interactions
cem <- plot_interactions(cem,n=30) # generate plot

pdf("figures/CEMiTool/network_hub_M1.pdf", width = 10, height = 7)
cem@interaction_plot$M1
dev.off()

pdf("figures/CEMiTool/network_hub_M5.pdf", width = 10, height = 7)
cem@interaction_plot$M5
dev.off()

pdf("figures/CEMiTool/network_hub_M2.pdf", width = 10, height = 7)
cem@interaction_plot$M2
dev.off()

pdf("figures/CEMiTool/network_hub_M3.pdf", width = 10, height = 7)
cem@interaction_plot$M3
dev.off()

pdf("figures/CEMiTool/network_hub_M4.pdf", width = 10, height = 7)
cem@interaction_plot$M4
dev.off()


write_csv(cem@module, "figures/CEMiTool/Gene_modules.csv")
```



